Combinatorics of the geometry of Wilson loop diagrams

### 2019-01-16 General WLD combinatorics.tex
Ammended theorem 2.24 to make clearer. Read through rest of 2.4.

\begin{proof}
One direction has been proved in previous work, but we give a different proof here to be consistent with the method of this document.

@@

Let $U \subset V(R)$ be any subset of size $|U| = |R|$. The set $U$ is independent in the subdiagram $(R, V(R))$, and thus in $W$. The complementary set $F(P)$ is a flat of maximal rank by lemma \ref{maxexactcomplementrank} ($\rk(F(P)) = |P|$). Let $B \subset F(P)$ be a maximal indpendent set ($\rk B = |P|$) in $F(P)$. Since $F(P)$ is a flat, adding any element of $V(R)$ to $B$ increases the rank. Therefore, any basis of $W$ can be written as $B \cup U$. However, since $F(P)$ is common to both $W$ and $W'$, and $V(R) = V(R')$, any any basis of $W'$ can also be written $B \cup U$. Thus both matroids have the same bases sets, proving that they are the same.

For the converse, assume that the matroids associated to $W$ and $W'$ are the same: $M(W) = M(W')= M$. Let $\{(P_i, V(P_i)\}_{i=1}^k$ and $\{(P'_i, V(P'_i)\}_{i=1}^l$ be the sets of maximally exact subdiagrams of $W$ and $W'$. Write $F_i = F(P_i)$ and $F'_i = F(P^{'c}_i)$ to be the complementary cyclic flats. By Lemma \ref{exactcircuitlem} $M/F_i$ is a uniform matroid. Thereofore, $k = l$, and we may write $V(P_i) = V(P'_i)$.

Reorganize the vertex sets as follows: \ba \cup_{P_i \neq P_i'} V(P_i) = V(\cup_{P_i \neq P'_i} P_i)  \label{vertexsets}\; .\ea Since maximal exact subdiagrams partition $\cP$, by Lemma \ref{uniqueproppartitioncor}, write $cup_{i = 1}^k P_i = \cup_{i = 1}^k P'_i = \cP$. Then equation \eqref{vertexsets} becomes \bas \cap_{P_i \neq P_i'} (F(P_i^c))^c = F(\cup_{P_j = P'_j} P_j)^c \;.\eas These flats may, of course, be empty.

Thus, we have partitioned the vertices of $W$ and $W'$ into a single set that supports a union of maximal exact subdiagrams whose propagators between $W$ and $W'$, and the complementary propagator flats that support the propagators in common between $W$ and $W'$. As mentioned above, the latter may be empty.

Define a family of Wilson loop diagrams, $W_0$ to $W_k$ defined such that $W_0 = W$ and $W_i$ is derived from $W_{i-1}$ by replacing the propagator set $P_i$ with $P'_i$. In this manner, $W' = W_k$ and $W_i \sim W_{i+1}$, making $W \sim W'$.
\end{proof}

Since there is a unique way to decompose $W$ into maximal exact subdiagrams, it is logical to ask how many diagrams there are in an equivalence class. It is a classical fact the the number of triangulations of an $n$-gon is the $n-2$ Catalan number, namely $\frac{1}{n-1}\binom{2(n-2)}{n-2}$.  Thus we can count the number of equivalent diagrams.

>>>>

\begin{proof}
One direction has been proved in \ref[Theorem 1.18]{wilsonloops}, but we give a different proof here to be consistent with the method of this document.

@@

Let $U \subset V(R)$ be any subset of size $|U| = |R|$. Since $(R, V(R))$ defines a uniform matroid, by Lemma \ref{exactuniformthm}, the set $U$ is independent in the subdiagram $(R, V(R))$, and thus in $W$. The complementary set $F(P)$ is a flat of maximal rank by lemma \ref{maxexactcomplementrank} ($\rk(F(P)) = |P|$). Let $B \subset F(P)$ be a maximal indpendent set ($\rk B = |P|$) in $F(P)$. Since $F(P)$ is a flat, adding any element of $V(R)$ to $B$ increases the rank. Therefore, any basis of $W$ can be written as $B \cup U$ for some $B$ and $U$ of the form indicated. However, since $F(P)$ is common to both $W$ and $W'$, and $V(R) = V(R')$, any any basis of $W'$ can also be written $B \cup U$. Thus both matroids have the same bases sets, proving that they are the same.

For the converse, assume that the matroids associated to $W$ and $W'$ are the same: $M(W) = M(W')= M$. Let $\mathcal{R} = \{(P_i, V(P_i)\}_{i=1}^k$ and $\mathcal{R}' = \{(P'_i, V(P'_i)\}_{i=1}^l$ be the sets of maximally exact subdiagrams of $W$ and $W'$. Write $F_i = F(P_i^c)$ and $F'_i = F(P^{'c}_i)$ to be the complementary cyclic flats. By Theorem \ref{exactuniformthm} and Theorem \ref{exact diagrams contractions},  $M/F_i$ and $M/F_i'$ are uniform matroids. Since, by Theorem \ref{exactuniformthm}, any subset $V$ of $M$ such that $M/(V^c)$ is uniform defines an exact subdiagram of both $W$ and $W'$. If $U \subset V$ are two such thsets, then the corresponding exact subdiagrams are also subsets. Therefore, $|\mathcal{R}| = |\mathcal{R}'|$, and we write $V(P_i) = V(P'_i)$, after possibly reindexing $\mathcal{R}$ and $\mathcal{R}'$.

Since the sets of propagators defining maximal exact subdiagrams partition $\cP$, by Lemma \ref{uniqueproppartitioncor}, write $\cup_{i = 1}^k P_i = \cup_{i = 1}^k P'_i = \cP$. Reorganize the vertex sets of maximal exact subdiagrams as follows: \ba \cup_{P_i \not \in \mathcal{R}'} V(P_i) = V(\cup_{P_i \not \in \mathcal{R}'} P_i)  = F(\cup_{P_i \in \mathcal{R}'} P_i)^c\label{vertexsets}\; .\ea The final flat may, of course, be empty.

Thus, we have partitioned the vertices of $W$ and $W'$ into two complementary sets. The first, $ \cup_{P_i \not \in \mathcal{R}'} V(P_i)$,  is comprised of the union of supports of maximal exact subdiagrams whose propagators differ between $W$ and $W'$. The second, $ F(\cup_{P_i \in \mathcal{R}'} P_i)$,  is the propagator flat of all propagators in common between $W$ and $W'$.

Without loss of generality, assume that $|\mathcal{R}| = |\mathcal{R}'| = 1$. Then the two Wilson loop diagrams are equivalent. If  $|\mathcal{R}| = |\mathcal{R}'| > 1$, then one may define a family of Wilson loop diagrams, $W_0$ to $W_k$ defined such that $W_0 = W$ and $W_i$ is derived from $W_{i-1}$ by replacing the propagator set $P_i$ with $P'_i$. In this manner, $W' = W_k$ and $W_i \sim W_{i+1}$, making $W \sim W'$.
\end{proof}

Since there is a unique way to decompose $W$ into maximal exact subdiagrams, it is logical to ask how many diagrams there are in an equivalence class. It is a classical fact the the number of triangulations of an $n$-gon is the $n-2$ Catalan number, namely $\frac{1}{n-1}\binom{2(n-2)}{n-2}$.  Thus we can count the number of equivalent diagrams. [CITATION?]
